Personal Total Income of Individual in the United States of America in 2023

### 2024-10-24 paper/paper.qmd
Abstract and Introduction updated

---
title: "Counts of Crimes and Their Subtypes in Toronto from 2014 to 2023"
subtitle: "Details and Trends"
author: 
  - Justin (Jiazhou) Bi
thanks: "Inspired and instructed by: https://github.com/RohanAlexander/marriage. All the project related files can be found at: https://github.com/Jiazhou-Bi/reported-crime/tree/master."
date: Sept. 23, 2024
date-format: long
abstract: "This report analyzes crime trends and subtypes in Toronto from 2014 to 2023, highlighting fluctuations and speculating possible socio-economic, technological, and pandemic-related factors influencing criminal activity. Using data from Toronto Open Data, the study aggregates and examines reported crimes in categories such as property crimes, violent offenses, and cybercrime. The results show a significant increase in reported crimes, from 113,426 in 2014 to 169,620 in 2023. A notable drop in crime during 2020 and 2021 corresponds with COVID-19 lockdowns, which reduced certain physical crimes while increasing cybercrimes like fraud. Despite the overall rise in crime, clearance rates for crimes like fraud and theft significantly declined, underscoring challenges for law enforcement."
format: pdf
jupyter: python3 
number-sections: true
bibliography: references.bib
---

@@

Crime is defined as "an action or omission which constitutes an offense and is punishable by law." (@TheOxfordDictionary_1992) However, this seemingly straightforward definition can misguide us to think that crime is a simple and clear concept with specific defining criteria, a swift process of judgment, and pre-defined punishment. Crime is a complete concept that affects every aspect of all individuals, communities, and societies. Crimes can reshape the power dynamics of our social structure, influence our public policies, and even change our day-to-day interactions with others. As a result, understanding the trends in crimes and the variations in their subtypes is crucial for policymakers, law enforcement, and every member of our society to devise better strategies to prevent crimes from happening. Therefore, this report aims to explore and examine the current of reported crimes in Toronto from 2014 to 2023, discover and highlight some fluctuations, and propose some potential underlying factors that might have contributed to these trends. The raw data being used in the report is available from Toronto Open Data. (@TorontoOpenData) For the data analysis, I have used pandas (@pandas). For data visualization, I have used Seaborn. (@seaborn) All the tables were generated by using Plotly. (@plotly)

This analysis can be helpful by providing valuable insights into current crime activity trends. Certain crimes, such as violent offenses, theft, and property crimes, are believed to be highly correlated with broader socio-economic conditions. (@lofstrom2016crime) Meanwhile, with new technology emerging, other types of criminal activity, such as technology-facilitated fraud, are becoming more prominent. According to RCMP's report, Canadian victims have lost 530 million dollars due to fraud. Thus, examining various types of crime can assist us to better understand these crimes and devise better strategies to better protect our safety and properties.

On the other hand, between 2014 and 2023, the world experienced significant transformations that may impact crime rates. In particular, the COVID-19 pandemic, which began in 2020 and ended in early 2022, has disrupted our daily routines and led to shifts in criminal behaviour. Lockdowns and other public health measures reduced opportunities for certain types of crime, such as burglaries and public disturbances, while simultaneously leading to increases in domestic violence and cybercrime. Additionally, economic hardship caused by the pandemic, including rising unemployment rates and financial instability, may have driven a surge in financially motivated crimes, such as fraud and theft.

This report will delve into crime statistics from 2014 to 2023 in Toronto. I will closely examine all the reported crime categories and specific subtypes of crimes. By analyzing these trends, I aim to clarify how crime has evolved over the last decade and identify possible socio-economic, technological, and legal factors that have driven these changes. Through this analysis, we can gain valuable insights into how best to address the challenges posed by both traditional and emerging forms of crime in the future.

@@

import pandas as pd
data = pd.read_csv('../data/analysis_data/analysis_data.csv', index_col=False)
data.index = data.index + 1
data.head()
# styled_data = data.head().style.format({'ClearRate': "{:.2f}"}).hide().set_table_styles([{"selector":"th.col_heading,td","props":[("width","100px")]}])
# styled_data
#data.values[:5]
#print(data.head().to_string(index=False))

@@

#| echo: false
#| label: tbl-table2
#| warning: false
#| tbl-cap: Six Crime Categories
import pandas as pd
import plotly.graph_objects as go
data = pd.read_csv('../data/analysis_data/analysis_data.csv')
unique_category = data.Category.unique()
category_df = pd.DataFrame(unique_category, columns=['Categories'])
category_df.index = category_df.index + 1
category_df
```

@@

```{python}
#| echo: false
#| label: tbl-table3
#| warning: false
#| tbl-cap: Crime Categories and Subtypes
import pandas as pd
data = pd.read_csv('../data/analysis_data/analysis_data.csv')
unique_category_subtype_data = data.drop_duplicates(subset=['Category', 'Subtype'])
unique_category_subtype_df = unique_category_subtype_data[['Category', 'Subtype']]
unique_category_subtype_df = unique_category_subtype_df.sort_values(by=['Category','Subtype'],ascending=True)
unique_category_subtype_df.reset_index(drop=True, inplace=True)
unique_category_subtype_df.index = unique_category_subtype_df.index + 1
unique_category_subtype_df
```

>>>>

---
title: "Personal Total Income of Individual in the United States of America in 2023"
subtitle: "A Predictive Model Using the American Community Survey Data"
author: 
  - Justin (Jiazhou) Bi and Weiyang Li
thanks: "Inspired and instructed by: https://github.com/RohanAlexander/marriage. All the project related files can be found at: https://github.com/Jiazhou-Bi/ACS-Income-Model/tree/main."
date: Oct 25, 2024
date-format: long
abstract: "This project aims to build a predictive model for personal total income using various demographic attributes, such as marital status, occupation, and residing state. We explored three popular machine learning algorithms: Random Forest, Linear Regression, and Extreme Gradient Boosting. Ultimately, Extreme Gradient Boosting was selected due to its superior predictive performance. While the model achieved a high R-squared value, it also exhibited a high Mean Squared Error, indicating challenges in accurately predicting all data points. Future steps for improving the model may involve incorporating additional demographic features to enhance model accuracy and reduce errors. Additionally, fine-tuning hyperparameters, experimenting with feature engineering, and handling outliers could further improve performance."

format: pdf
jupyter: python3 
number-sections: true
bibliography: references.bib
---

@@

Income is a fundamental concept as it affects nearly every member of society, influencing access to resources, quality of life, and overall economic stability. At the same time, it is also a highly complex concept to define, as it can be interpreted differently depending on various perspectives, such as economic, social, legal, and tax purposes. Therefore, we must examine various demographical factors that may impact income to understand its concept better. Our project defines income as "all forms of net financial resources generated or lost from work, investments, or other activities." This project aims to predict personal income using data from the 2023 American Community Survey (ACS) (@ipums_usa_v15). The ACS provides detailed demographic information, including age, education, employment, marital status, mortgage status, veteran status, etc... As a result, it provides a robust dataset for studying potential factors that may influence income.

Due to the timeframe limitations, the findings of this project are specifically applicable to the 2023 dataset. As such, the results may vary across different periods, and their generalizability has not been assessed within this project. The raw data used in this analysis was sourced from IPUMS using the IPUMS API. (@ipums_api) We utilized pandas (@pandas) for data processing, and data visualization was carried out using Seaborn. (@seaborn) Additionally, all tables presented in this report were generated with Plotly. (@plotly)

Income inequality is a serious problem and can lead to significant consequences and it may affect social structure, economic stability, as well as policy-making processes. Researchers found that demographic factors, such as education level and occupation type, can significantly impact an individual's financial standing. (@Autor_2013) Similarly, social relationships, such as marital status and household characteristics, affect income dynamics. (@WaiteGallagher_2000) Therefore, applying machine learning models to analyze the ACS dataset and using the information mentioned to predict personal income is reasonable.

We have chosen 2023 ACS dataseet because it offers many demographical details, as well as self-reported total income of the individuals. The sample is also representative and covers all geographical locations in the USA. Furthermore, it provides easy access to their database through API. We will employ advanced predictive models, including random forest, lineaer regression, and extreme gradient boosting machine to estimate personal income based on demographic factors to provide insights into which potential factors have the most significant impact on personal income. We hope the findings generated from our project can provide insights to policy-making process so that we can, hopefully, reduce income inequality and support further economic development in the future.

If we can successfully predict personal income using demographic information from the survey, this analysis will be able to contribute to a nuanced understanding of the American economy in 2023. It can help to reveal how demographic changess, educational attainment, and employment status/type can impact personal income. On the other hand, if the model cannot predict personal income with a high degree of confidence, this would suggest that the factors included in the analysiscannot not fully capture the intricacies of economic behavior and individual circumstances. As a result, additional variables, such as detailed education background like school of graduation, languages spoken, health status, etc, might be necessary in accurately predicting income.

@@

import pandas as pd
data = pd.read_parquet('../data/02-analysis_data/cleaned_data.parquet')
data.index = data.index + 1
data.head()
# styled_data = data.head().style.format({'ClearRate': "{:.2f}"}).hide().set_table_styles([{"selector":"th.col_heading,td","props":[("width","100px")]}])
# styled_data
#data.values[:5]
#print(data.head().to_string(index=False))

@@

#| echo: false
#| label: tbl-table2
#| warning: false
#| tbl-cap: Six Crime Categories
# import pandas as pd
# import plotly.graph_objects as go
# data = pd.read_csv('../data/analysis_data/analysis_data.csv')
# unique_category = data.Category.unique()
# category_df = pd.DataFrame(unique_category, columns=['Categories'])
# category_df.index = category_df.index + 1
# category_df
```

@@

```{python}
#| echo: false
#| label: tbl-table3
#| warning: false
#| tbl-cap: Crime Categories and Subtypes
# import pandas as pd
# data = pd.read_csv('../data/analysis_data/analysis_data.csv')
# unique_category_subtype_data = data.drop_duplicates(subset=['Category', 'Subtype'])
# unique_category_subtype_df = unique_category_subtype_data[['Category', 'Subtype']]
# unique_category_subtype_df = unique_category_subtype_df.sort_values(by=['Category','Subtype'],ascending=True)
# unique_category_subtype_df.reset_index(drop=True, inplace=True)
# unique_category_subtype_df.index = unique_category_subtype_df.index + 1
# unique_category_subtype_df
```
